---
title: "bioRxivGrowth"
author: "Chinmay Shukla"
date: "1/3/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in input data

To begin our analysis, we will first read in data about number of papers submitted to bioRxiv every month. I compiled this data manually from http://www.biorxiv.org/archive

```{r bioRxiv}
bioRxiv <- read.table("../data/bioRxivGrowth.tsv", sep="\t", header=T)
```

Before analyzing the data, lets see the type of the bioRxiv object which we used to read data from the input file. 

```{r dataStructure}
str(bioRxiv)
typeof(bioRxiv$Month)
```

We see that bioRxiv is a data frame of integers. While this works fine for Preprints, Month and perhaps even year should be stored as factors. Another problem is that the default ordering of defining factors would not work very well here and so we should tell R exactly how we want the months arranged. Finally, while abbreviations are used for most months, March and April should be abbreviated for consistency. Such problems are common in real world data sources and hence I decided to include this here. 

So now lets do a little data wrangling.

```{r dataWrangling}
bioRxiv$Month <- gsub("March", "Mar", bioRxiv$Month)
bioRxiv$Month <- gsub("April", "Apr", bioRxiv$Month)
bioRxiv$Month_f <- factor(bioRxiv$Month, 
                          levels=c("Jan", "Feb", "Mar", "Apr", "May", "June", "July",
                                   "Aug", "Sep", "Oct", "Nov", "Dec"))
```

## Exploratory Plots

We can start to plot the data and see if we see any interesting trends now. I am going to make a bar graph of the # of Preprints in each month and fragment it by Year. 

```{r eda}
library(ggplot2)
library(extrafont)
gg <- ggplot(bioRxiv, aes(x=Month_f, y=Preprints)) + geom_bar(stat="identity") + facet_grid(.~Year) + theme_bw()
```

We can use the steps described in our post on prettifying ggplots to make this look more professional. Below I have described the steps as a function.

```{r ggplotPretty}
prettyGG <- function(gg, xLab, yLab, titleLab, subtitleLab, 
                     fontFamily="Verdana", axisTextSize=8, xJust, yJust,
                     axisTitleSize=11, plotTitleSize=14, plotSubtitleSize=11)
  {
  ggPretty <- gg + labs(x=xLab, y=yLab, title=titleLab, subtitle = subtitleLab)
  ggPretty <- ggPretty + theme_bw(base_family = fontFamily)
  ggPretty <- ggPretty + theme(plot.margin=margin(10, 10, 10, 10))
  ggPretty <- ggPretty + theme(panel.grid.minor=element_blank())
  ggPretty <- ggPretty + theme(panel.grid.major=element_line(linetype="dotted", size=0.5))
  ggPretty <- ggPretty + theme(axis.ticks=element_blank())
  ggPretty <- ggPretty + theme(axis.text=element_text(size=axisTextSize))
  ggPretty <- ggPretty + theme(axis.text.x=element_text(hjust=xJust))
  ggPretty <- ggPretty + theme(axis.text.y=element_text(vjust=yJust))
  ggPretty <- ggPretty + theme(axis.title.x=element_text(hjust=0.5, face="bold", size=axisTitleSize))
  ggPretty <- ggPretty + theme(axis.title.y=element_text(vjust=0.5, face="bold", size=axisTitleSize))
  ggPretty <- ggPretty + theme(plot.title=element_text(hjust=0, face="bold", size=plotTitleSize))
  ggPretty <- ggPretty + theme(plot.subtitle=element_text(hjust=0, face="italic", size=plotSubtitleSize))
  
  return(ggPretty)
}
```


```{r PrettyPlot, out.width='2000px', dpi=600}
ggPretty <- prettyGG(gg, xLab="Month", yLab = "# of Preprints", 
                     titleLab = "Number of preprints submitted to bioRXiv every month",
                     subtitleLab = "The number of preprints submitted to bioRxiv from its inception in Nov 2013 has increased exponentially", 
                     fontFamily = "Verdana", axisTextSize=4,
                     xJust = c(rep(0.5,12)), yJust = c(0, 0.5, 0.5, 1), 
                     axisTitleSize=8, plotTitleSize=11, plotSubtitleSize=8)

ggPretty

```

Another way to summarize and understand the explosion in preprints on bioRxiv is to look at how the median number of preprints in a month has changed from 2013 to 2016. Lets do that now. We will use dplyr to group our data by year and summarize using the median number of preprints in each group (year) and like always ggplot to display our results.

```{r medianMonthlyPreprints, out.width='2000px', dpi=600}
library(dplyr)
byYear <- group_by(bioRxiv, Year)
byYear <- summarise(byYear, median(Preprints))
colnames(byYear) <- c("Year", "medianPreprints")
gg <- ggplot(byYear, aes(x=factor(Year), y=medianPreprints)) + geom_bar(stat="identity") + theme_bw()
ggPretty <- prettyGG(gg, xLab = "Year", 
                     yLab = "Median monthly Preprints submitted to bioRxiv", 
                     titleLab = "Median number of preprints submitted to bioRxiv in a month for each year", 
                     subtitleLab = "Every year the median number of articles submitted to bioRxiv in a month increases by a factor of 2 or more", 
                     axisTextSize = 6, 
                     xJust = c(0, rep(0.5, 2), 1), yJust = c(0, rep(0.5, 3), 1),
                     axisTitleSize = 8, plotTitleSize = 10, plotSubtitleSize = 8)
ggPretty
```

One of the things I am curious about is if the number of preprints submitted to bioRxiv peaks during a particular month. One simple way to answer this question is to divide our data by year and then look at the months with the 3 highest submissions. Since bioRxiv came online only in Nov 2013, we will discard 2013 from this analysis

```{r maxMonthByYear}
rXiv2014 <- bioRxiv %>% filter(Year=="2014") %>% arrange(desc(Preprints))
rXiv2015 <- bioRxiv %>% filter(Year=="2015") %>% arrange(desc(Preprints))
rXiv2016 <- bioRxiv %>% filter(Year=="2016") %>% arrange(desc(Preprints))

rXiv2014$Month[1:3]
rXiv2015$Month[1:3]
rXiv2016$Month[1:3]
```

This is very interesting. It seems the preprints submitted peak towards the end of the year. There could be 2 factors - 1). Scientists are rushing to get their papers out in a particular calendar year and Dec is the deadline for that and so there are a lot of submissions then or 2).bioRxiv is continuously increasing in its popularity and so as time goes by more people are submitting their research. Hence for a given year, the peak would be at the end of year.

My guess is that its a combination of the 2 but time will help us know if one factor is particularly dominant over the other.