---
title: "eplWinner"
author: "Chinmay Shukla"
date: "12/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjson)
library(tidyverse)

```

### Load data from R object.

Previously, I downloaded JSON data from a few sources (ESPN API mainly), cleaned it up and made a EPL object from it. It has been saved here. The first step of the analysis is to load this object (a tibbled called EPL).

```{r loadData, cache=FALSE, message=FALSE}
readRDS('~/Documents/Research/OTSD/data/epl.rds')
```

### On average, how many games does the EPL champion win?

```{r gamesWon, cache=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(filter(epl, rank%in%c(1,2,3,4)), aes(x=1, y=prcntWins, color=rank)) + geom_boxplot() + theme_bw() + 
  xlab('') + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = c(0.9,0.8)) + 
  scale_y_continuous(limits=c(25,80))

ggplot(filter(epl, rank%in%c(1,2,3,4)), aes(x=prcntPts, color=rank)) + 
  geom_density() + theme_bw() + theme(legend.position = c(0.9,0.8))
```

### On average, how many games does the EPL champion draw?

```{r gamesDrawn, cache=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(filter(epl, rank%in%c(1,2,3,4)), aes(x=1, y=prcntDraws, color=rank)) + geom_boxplot() + theme_bw() + 
  xlab('') + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = c(0.9,0.8)) 

ggplot(filter(epl, rank%in%c(1,2,3,4)), aes(x=prcntDraws, color=rank)) + 
  geom_density() + theme_bw() + theme(legend.position = c(0.9,0.8))
```

## On average, how many games does the EPL champion draw?

```{r gamesLost, cache=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(filter(epl, rank%in%c(1,2,3,4)), aes(x=1, y=prcntLosses, color=rank)) + geom_boxplot() + theme_bw() + 
  xlab('') + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = c(0.1,0.8)) 

ggplot(filter(epl, rank%in%c(1,2,3,4)), aes(x=prcntLosses, color=rank)) + 
  geom_density() + theme_bw() + theme(legend.position = c(0.9,0.8))
```

